---
title: "Regional Data Profiles"
output: 
  html_document:
    theme: cerulean
    toc: false
    toc_float: false
    number_sections: false
---

This code pulls data from the Central Database (aka Elmer) and creates graphics and tables for the Regional Data Profile Pages.
```{r setup, include=FALSE}

# Data Processing and Wrangling
library(dplyr)
library(lubridate)

# Database Connection
library(odbc)
library(DBI)

# Plotting Libraries
library(ggplot2)
library(scales)
library(plotly)

# Custom Colors
psrc_colors <- c(
  "CoastRhodo" = "#91268F",
  "CedarShake" = "#F05A28",
  "DouglasFirShoot" = "#8CC63E",    
  "FerryWake" = "#00A7A0",
  "DarkGrey" = "#76787A",    
  "LightGrey" = "#BBBDC0",
  "King County" = "#AD5CAB",
  "Kitsap County" = "#F4835E",
  "Pierce County" = "#A9D46E",
  "Snohomish County" = "#40BDB8",
  "Historical" = "#76787A",
  "Forecast" = "#AD5CAB"
)

pop_colors <- c(
  "Historical" = "#76787A",
  "Forecast" = "#AD5CAB"
)

hh_colors <- c(
  "Historical" = "#76787A",
  "Forecast" = "#F4835E"
)

# SQL Database Connection settings
elmer_connection <- dbConnect(odbc::odbc(),
  driver = "SQL Server",
  server = "AWS-PROD-SQL\\Sockeye",
  database = "Elmer",
  trusted_connection = "yes"
  )

# Macro-forecast Tables
macro_pop_facts<- "Macroeconomic.pop_facts"
macro_pop_dims<- "Macroeconomic.pop_group_dims"

macro_hh_facts<- "Macroeconomic.household_facts"

initial_forecast_year <- 2020

```

``` {r useful_functions, include = FALSE}

create_plot <- function(w_data, w_fill, w_y, w_x, bar_type, label_type, w_colors) {
  w_chart <- ggplot(data=w_data, aes(fill=get(w_fill), y=get(w_y), x=get(w_x))) +
                    geom_bar(position=bar_type, stat="identity") +
                    scale_fill_manual(values= w_colors) +
                    scale_y_continuous(labels = label_type) +
                    theme_light() +
                      theme(
                        axis.text=element_text(size=10),
                        axis.text.x.bottom=element_text(size=10),
                        axis.title.y = element_blank(),
                        axis.title.x = element_blank(),
                        panel.grid.major = element_line(colour="#BBBDC0",size = 0.25),
                        panel.grid.minor = element_line(colour="#BBBDC0",size = 0.25),
                        panel.border = element_blank(),
                        axis.line = element_blank(),
                        legend.position="bottom",
                        legend.title = element_blank())
  return(w_chart)
}

```

``` {r macro_forecast_data, include = FALSE}

forecast_years <- c(2010,2015,2020,2025,2030,2035,2040,2045,2050)

# Pull data from Elmer and store as a tibble for use in dplyr
forecasted_population_data <- as_tibble(dbReadTable(elmer_connection,SQL(macro_pop_facts)))
forecasted_population_dims <- as_tibble(dbReadTable(elmer_connection,SQL(macro_pop_dims)))
forecasted_household_data <- as_tibble(dbReadTable(elmer_connection,SQL(macro_hh_facts)))

# Add Age Group Detail to the Population Table and trim down the final table
forecasted_population_data <- inner_join(forecasted_population_data, forecasted_population_dims, by="pop_group_dim_id") %>%
  select(data_year,population,group_name)

total_population <- forecasted_population_data %>%
  filter(group_name == "Total") %>%
  filter(data_year %in% forecast_years) %>%
  mutate(group_name=replace(group_name, forecast_years<initial_forecast_year, "Historical")) %>%
  mutate(group_name=replace(group_name, forecast_years>=initial_forecast_year, "Forecast"))

total_households <- forecasted_household_data %>%
  select(data_year,households,dataset_dim_id) %>%
  filter(data_year %in% forecast_years) %>%
  mutate(dataset_dim_id=replace(dataset_dim_id, forecast_years<initial_forecast_year, "Historical")) %>%
  mutate(dataset_dim_id=replace(dataset_dim_id, forecast_years>=initial_forecast_year, "Forecast"))

# Remove objects that are no longer needed from memory
rm("forecasted_population_dims") 

```

# Forecast Population & Households {.tabset .tabset.fade .tabset-pills}
PSRC forecasts the region's households, persons, jobs, and other economic and demographic variables through the year 2050.

The Macroeconomic Forecast is an input to PSRC's land use and travel forecasting, and provides the growth assumptions used in our regional growth strategy. It is developed by PSRC with help and review by technical staff from state and local government. The Macroeconomic Forecast is the resource behind the forecasts of regional population and households shown below. For more information on the Macroeconomic Forecst, please click [here](https://www.psrc.org/regional-macroeconomic-forecast).

## Population

``` {r pop_forecast_chart, echo = FALSE}

pop_chart <- create_plot(total_population, "group_name", "population", "data_year", "dodge", scales::comma, pop_colors)
pop_chart

```

## Households

``` {r hh_forecast_chart, echo = FALSE}

hh_chart <- create_plot(total_households, "dataset_dim_id", "households", "data_year", "dodge", scales::comma, hh_colors)
hh_chart

```

``` {r final_cleanup, include = FALSE}

# Disconnect from Elmer
odbc::dbDisconnect(elmer_connection)

```
